opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-12-04 chapters/01_introduction.tex
Merged Xqueyer paragraph into contribution entry

Additionally, attempting to refine all crystal structure parameters at once, starting from this initial structure, is known to lead to unphysical results\cite{Ozaki2020}. Hence parameters are refined iteratively, with each iteration only refining a limited set of parameters. If, during the iterative determination of parameters, the structural model becomes unphysical the operator must intervene to backtrack to a previous successful step or start over entirely. Finding the correct order in which to refine structure parameters and finding the correct values for fixed parameters both present problems that add to the difficulty of the refinement process.

%What makes the refinement process time-consuming and difficult is finding the correct order in which to refine the crystal structure parameters\cite{Ozaki2020}. \\ % \\

@@

% Predicting structure information with machine learning
Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic, Wang2020}
The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions. The physical effects underlying these imperfections need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms. Experimental data can provide clues on how to model these imperfections in a more realistic fashion. \\

At the moment, the number of openly available datasets for experimental powder data is very limited. The main sources for experimental patterns are XXX, YYY, and ZZZ. Further commercial datasets such as KKK and LLL exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability. Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is currently out of reach. The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms,
%For example 4M sim patterns\cite{cao2024simxrd} = 4 Mio. patterns
which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here. \\

>>>>

Additionally, attempting to refine all crystal structure parameters at once, starting from this initial structure, is known to lead to unphysical results\cite{Ozaki2020}. Hence parameters are refined iteratively, with each iteration only refining a limited set of parameters. If, during the iterative determination of parameters, the structural model becomes unphysical the operator must intervene to backtrack to a previous successful step or start over entirely. Finding the correct order in which to refine structure parameters and finding the correct values for fixed parameters both present problems that add to the difficulty of the refinement process.\\

@@

% Predicting structure information with machine learning
Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004, zhang2024crystallographic}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021,zhang2024crystallographic}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023,zhang2024crystallographic, Wang2020, Vecsei2018} The patterns observed in experiments exhibit imperfections that do not appear when simulating the X-ray diffraction pattern of a powder sample under ideal conditions. The physical effects underlying these imperfections need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms. Experimental data can provide clues on how to model these imperfections in a more realistic fashion. \\

At the moment, the number of openly available datasets for experimental powder data is very limited. The main sources for experimental patterns are XXX, YYY, and ZZZ. Further commercial datasets such as KKK and LLL exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability. Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is currently out of reach. The most advanced ML models currently are trained on approximately $10^5 - 10^6$ simulated diffractograms. \cite{Salgado2023, Schopmans2023}
This is - to the best of our knowledge - two orders of magnitude larger than the largest currently curated experimental dataset (PDF-5+ with approximately 20,000 experimental patterns) and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here. \\

### 2024-12-04 chapters/02_SOTA_ML.tex
Merged Xqueyer paragraph into contribution entry

It becomes apparent, that the more general the task is, the more challenging the transfer to experimental data is. For example, the space group classification task is very general, for all material systems. Therefore, transferring it to the application on experimental diffraction patterns is difficult. \cite{Schopmans2023, Lee2022, Park2017} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex. \cite{Schuetzke2024, Lee2020}

The ultimate goal of ML efforts is to make PXRD data identification practical for experimental use, automating structure property prediction. However, the current volume of experimental PXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental PXRD database. To address this, two key approaches are essential. First, developing more sophisticated simulation methods to better approximate experimental patterns\cite{cao2024simxrd} by using statistics from experimental diffractograms. Second, creating an experimental database that enables transfer learning to bridge the gap between simulated and real-world data. For both of these steps, the development of opXRD is particularly significant, as it will provide a comprehensive experimental benchmark for the community, allowing fair comparison of baseline models and accurate evaluation of their applicability in real experimental situations.

>>>>

It becomes apparent, that the more general the task is, the more challenging the transfer to experimental data is. For example, the space group classification task is very general, for all material systems. Therefore, transferring it to the application on experimental diffraction patterns is difficult. \cite{Schopmans2023, Lee2022, Vecsei2018} On the other hand, there are some successful approaches that also work well on experimental data, but those are mostly methods that do phase determination in a limited compound space, making the task less complex. \cite{Schuetzke2024, Lee2020}

The ultimate goal of ML efforts is to make pXRD data identification practical for experimental use, automating structure property prediction. However, the current volume of experimental pXRD patterns is insufficient to effectively train ML models, highlighting an urgent need for a comprehensive experimental pXRD database. To address this, two key approaches are essential. First, developing more sophisticated simulation methods to better approximate experimental patterns\cite{cao2024simxrd} by using statistics from experimental diffractograms. Second, creating an experimental database that enables transfer learning to bridge the gap between simulated and real-world data. For both of these steps, the development of opXRD is particularly significant, as it will provide a comprehensive experimental benchmark for the community, allowing fair comparison of baseline models and accurate evaluation of their applicability in real experimental situations.

### 2024-12-04 chapters/04_our_dataset.tex
Described python + zenodo data availability
Merged Xqueyer paragraph into contribution entry

We extracted experimental pXRD data from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}. The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries. The data in the COD is placed in the public domain and licensed under the CC0 License. Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies. These 5432 structures only account for 1\% of the total COD database, but this is to be expected since most crystal structures are resolved from single-crystal diffraction. Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself. We could extract raw experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data. \\

@@

In the past two years, we have established a small-scale experimental powder X-ray diffraction database called the X-Ray Phase Identification Public Experimental Dataset (XRed). \footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations. \\

The data in XRed were generated in our laboratory using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources. Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters. These patterns are labeled with CIF files that document the refined structures. The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks. This project is progressing toward integration with the comprehensive opXRD database to create a large-scale, long-term experimental resource.

>>>>

Experimental pXRD data was extracted from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}.

%We extracted experimental pXRD data from the Crystallography Open Database .

The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries. The data in the COD is placed in the public domain and licensed under the CC0 License. Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies. These 5432 structures only account for 1\% of the total COD database, but this is to be expected since most crystal structures are resolved from single-crystal diffraction. Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself. We could extract raw experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data. \\

@@

A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations. \\

The data in XRed were generated in our laboratory using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources. Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters. These patterns are labeled with CIF files that document the refined structures. The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks. \\

Additionally, Xqueryer,\footnote{Xqueryer can be accessed at \url{http://xqueryer.caobin.asia/}.} a cutting-edge free online platform for intelligent structure identification has been launched. Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database. Uploaded data are securely stored and integrated into the comprehensive opXRD database to progress toward a large-scale, long-term experimental resource. All usage complies with Xqueryer's terms of service.

### 2024-12-04 chapters/05_how_to_use.tex
Merged Xqueyer paragraph into contribution entry

The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14254271}.} and can be downloaded by any user without any barriers or restrictions. The database comes in two files, \"opxrd.zip\" and \"opxrd\_in\_situ.zip\". The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements. Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series. \\

Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library \"opxrd\" to easily download and interface with the dataset. The instructions for how to install this library can be found in the repository associated with the library.\footnote{The repository to this library can is located at \url{https://github.com/aimat-lab/opxrd}.} The opxrd library allows the dataset to be accessed using one simple command \pyth{OpXRD.load(root_dirpath)}. If the database is locally available under \pyth{root_dirpath} this command loads the library from this location. If the database is not available locally at this location, the database is automatically downloaded from Zenodo to \pyth{root_dirpath}. \\

>>>>

The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14254271}.} and can be downloaded by any user without any barriers or restrictions. The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''. The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements. Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series. \\

Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library \"opxrd\" to easily download and interface with the dataset. The instructions for how to install this library can be found in the repository associated with the library.\footnote{The repository to this library can is located at \url{https://github.com/aimat-lab/opxrd}.} The opxrd library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}. If the database is locally available under \pyth{root_dirpath} this command loads the library from this location. If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}. \\
